---
title: "NRG1 vs Scramble Proteomics Analysis (24h)"
author: "Marc Cantero"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    theme: journal
    highlight: kate
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# Set your working directory explicitly - change this path to your project folder
project_dir <- getwd()  # Or use an explicit path like: "/Users/marc/Projects/NRG1_Proteomics"
knitr::opts_knit$set(root.dir = project_dir)
setwd(project_dir)

cat("Working directory set to:", project_dir, "\n")
```

### Loading Required Packages

```{r load_packages, include=FALSE, echo=FALSE, message=FALSE}
suppressPackageStartupMessages({
  library(readxl)
  library(limma)
  library(showtext)
  library(viridis)
  library(dplyr)
  library(ggplot2)
  library(ggrepel)
  library(reshape2)
  library(tidyr)
  library(stringr)
  library(kableExtra)
  library(readr)
  library(pheatmap)
  library(Biobase)
  library(tidyverse)
  library(gprofiler2)
  library(ComplexHeatmap)
  library(clusterProfiler)
  library(org.Hs.eg.db)
  library(GO.db)
  library(enrichplot)
  library(VennDiagram)
  library(sysfonts)
  library(rcartocolor)
})
```

### Setting Up Visuals

```{r HERD_visual}
herd_palette <- c(
  "#14767F",  # RENEW Teal (dark teal)

  "#D55E00",  # Oxygenated Red (orange-red)
  "#117733",  # Leaf Green (dark green)
  "#882255",  # Deep Burgundy (dark purple-red)
  "#0072B2",  # Primary Blue (medium blue)
  "#E69F00",  # Bright Orange (gold-orange)
  "#6BC3B3",  # RENEW Aqua (light teal)
  "#2E0854",  # Indigo (very dark purple)
  "#44AA44",  # Positive Green (bright green)
  "#722F37",  # Wine (dark red)
  "#87CEEB",  # Sky Blue (light blue)
  "#F1C232",  # Light Gold (yellow-gold)
  "#50C878",  # Emerald (medium green)
  "#CC79A7",  # Cool Purple (pink-purple)
  "#001F3F",  # Navy (very dark blue)
  "#CCFF00",  # Lime (bright yellow-green)
  "#FFB6C1",  # Hot Pink (light pink)
  "#4A2511",  # Chocolate (dark brown)
  "#9370DB",  # Violet (medium purple)
  "#F4E4C1",  # Sand (light beige)
  "#FF4444",  # Vermilion (bright red)
  "#00D9A3",  # Bluish Green (bright cyan)
  "#FFFF00",  # Yellow (pure yellow)
  "#999999",  # Slate Grey (medium grey)
  "#E6E6FA",  # Lavender (very light purple)
  "#000000"   # Black
)

# Custom scale functions
scale_fill_herd <- function(values = NULL, n_colors = NULL, ...) {
  scale_fill_manual(
    values = values %||% herd_palette[seq_len(n_colors %||% length(herd_palette))],
    ...
  )
}

scale_color_herd <- function(values = NULL, n_colors = NULL, ...) {
  scale_color_manual(
    values = values %||% herd_palette[seq_len(n_colors %||% length(herd_palette))],
    ...
  )
}

# Treatment colors for your comparison
treatment_colors <- c(
  "Scramble" = "#999999",   # Control (Healthy)
  "NRG1" = "#9370DB"        # Treatment (Disease)
)
```

### Loading and Preprocessing the Dataset

```{r load_proteomics_data}
# Read the Excel file - adjust path as needed
proteomics_data <- General_proteome_WithIBAQ_0_24

# View structure
cat("Dataset dimensions:", dim(proteomics_data), "\n")
cat("Column names:\n")
print(colnames(proteomics_data))
```

```{r preprocessing}
# Create a clean copy
proteomics_clean <- proteomics_data

# Remove Fasta.headers and iBAQ.peptides columns (metadata, not needed for LFQ analysis)
proteomics_clean <- proteomics_clean[, !names(proteomics_clean) %in% c("Fasta.headers", "iBAQ.peptides")]

# Identify LFQ columns
lfq_cols <- grep("^LFQ\\.intensity\\.", names(proteomics_clean), value = TRUE)
cat("LFQ columns found:\n")
print(lfq_cols)

# Create proteome dataframe with necessary columns
proteome <- proteomics_clean[, c("Protein.IDs", "Majority.protein.IDs", 
                                  "Protein.names", "Gene.names", 
                                  lfq_cols)]

# Clean column names - remove LFQ.intensity. prefix
colnames(proteome) <- gsub("^LFQ\\.intensity\\.", "", colnames(proteome))

# Clean up empty gene names
proteome$Gene.names[proteome$Gene.names == ""] <- NA

# View the cleaned data
cat("\nProteome dataset dimensions:", dim(proteome), "\n")
print(head(proteome))
```

```{r adding_single_uniprot_name, warning=FALSE}
# Extract the first Uniprot ID from Majority.protein.IDs column
proteome$UniprotID <- sapply(strsplit(as.character(proteome$Majority.protein.IDs), ";"), function(x) x[1])

# Reorder columns to put UniprotID after Gene.names
gene_names_pos <- which(colnames(proteome) == "Gene.names")
other_cols <- setdiff(colnames(proteome), "UniprotID")
col_order <- c(
  other_cols[1:gene_names_pos],
  "UniprotID",
  other_cols[(gene_names_pos + 1):length(other_cols)]
)

proteome <- proteome[, col_order]
print(head(proteome))
```

### Quality Control

```{r targets}
# Define sample information
# Your samples: NRG124h_1_Disease, NRG124h_2_Disease, NRG124h_3_Disease, 
#               Scramble_1_Healthy, Scramble_2_Healthy, Scramble_3_Healthy

sample_cols <- colnames(proteome)[6:11]  # LFQ intensity columns

targets <- data.frame(Sample = sample_cols)

# Extract treatment from sample names (match NRG1 anywhere in name)
targets$Treatment <- ifelse(grepl("NRG1", targets$Sample), "NRG1", "Scramble")

# Extract replicate number
targets$Rep <- gsub(".*_(\\d+)_.*", "\\1", targets$Sample)

# Extract condition (Disease/Healthy)
targets$Condition <- ifelse(grepl("Disease", targets$Sample), "Disease", "Healthy")

targets %>%
  kable(caption = "Sample Information",
        col.names = c("Sample", "Treatment", "Replicate", "Condition")) %>% 
  kable_styling(bootstrap_options = c("striped", "bordered", "hover"),
                full_width = FALSE,
                position = "center") %>%
  row_spec(0, bold = TRUE, color = "white", background = "#14767F") %>%
  column_spec(1:4, extra_css = "font-family: 'Helvetica', sans-serif; padding: 10px;")
```

```{r create_output_dirs}
# Create results directories
if (!dir.exists("results/qc")) {
  dir.create("results/qc", recursive = TRUE)
}
if (!dir.exists("results/differential")) {
  dir.create("results/differential", recursive = TRUE)
}
if (!dir.exists("results/volcano")) {
  dir.create("results/volcano", recursive = TRUE)
}
```

```{r non-zero_proteins, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
# Replace NAs with 0 for counting (only in LFQ columns)
proteome_PI <- proteome
proteome_PI[sample_cols][is.na(proteome_PI[sample_cols])] <- 0

# Count proteins with intensity > 0 per sample
tmp <- data.frame(
  Sample = sample_cols, 
  Total_proteins = colSums(proteome_PI[sample_cols] > 0)
)

# Match samples to treatments
tmp$Group <- targets$Treatment[match(tmp$Sample, targets$Sample)]

# Create the plot
p <- ggplot(tmp, aes(x = Sample, y = Total_proteins, fill = Group)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = treatment_colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, max(tmp$Total_proteins) * 1.05)) +
  labs(title = "Proteins with intensity > 0",
       x = "Sample",
       y = "Number of Proteins",
       fill = "Treatment") +
  theme_minimal() +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1, vjust = 1, size = 10),
    axis.text.y = element_text(size = 10),
    axis.title = element_text(size = 12, face = "bold"),
    plot.title = element_text(hjust = 0.5, size = 14, face = "bold"),
    legend.position = "right",
    legend.title = element_text(face = "bold"),
    panel.grid.major.x = element_blank(),
    panel.grid.minor = element_blank()
  )

ggsave("results/qc/proteins_intensity_plot.svg", plot = p, width = 10, height = 6, device = "svg")
ggsave("results/qc/proteins_intensity_plot.jpeg", plot = p, width = 10, height = 6, device = "jpeg", dpi = 300)
print(p)
```

```{r total_library_size, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
tmp <- data.frame(
  Sample = sample_cols, 
  Total_intensities = colSums(proteome[sample_cols], na.rm = TRUE)
)
tmp$Group <- targets$Treatment[match(tmp$Sample, targets$Sample)]
tmp <- tmp[order(tmp$Group), ]
tmp$Sample <- factor(tmp$Sample, levels = tmp$Sample)

# Create the plot
p <- ggplot(tmp, aes(x = Sample, y = Total_intensities, fill = Group)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = treatment_colors) +
  scale_y_continuous(expand = c(0, 0)) +
  labs(title = "Library size",
       x = "Sample",
       y = "Total Intensities",
       fill = "Treatment") +
  theme_classic() + 
  theme(
    axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90), 
    plot.title = element_text(hjust = 0.5)
  )

ggsave("results/qc/Total_Library_Size.svg", plot = p, width = 10, height = 6, device = "svg")
ggsave("results/qc/Total_Library_Size.jpeg", plot = p, width = 10, height = 6, device = "jpeg", dpi = 300)
print(p)
```

```{r log2_plot, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
par(bty = "L")
group_factor <- factor(targets$Treatment[match(sample_cols, targets$Sample)])
plotDensities(log2(proteome[sample_cols]), 
              group = group_factor,
              col = treatment_colors[levels(group_factor)],
              legend = "topright",
              main = "Log2 Intensity Distribution (Raw)")

# Save to results folder
svg("results/qc/Density_Plot_Raw.svg", width = 10, height = 6)
par(bty = "L")
plotDensities(log2(proteome[sample_cols]), 
              group = group_factor,
              col = treatment_colors[levels(group_factor)],
              legend = "topright",
              main = "Log2 Intensity Distribution (Raw)")
invisible(dev.off())

jpeg("results/qc/Density_Plot_Raw.jpeg", width = 10, height = 6, units = "in", res = 300)
par(bty = "L")
plotDensities(log2(proteome[sample_cols]), 
              group = group_factor,
              col = treatment_colors[levels(group_factor)],
              legend = "topright",
              main = "Log2 Intensity Distribution (Raw)")
invisible(dev.off())
```

### Preprocessing

Data pre-processing workflow:

1. **Filter by Presence**: Protein should be present in at least 2 out of 3 replicates in at least one condition
2. **Log2-transformation**: Variance stabilization
3. **Median Centering**: Normalization to address multiplicative noise

```{r elist_creation}
# Set row names using Gene.names and UniprotID
rownames(proteome) <- paste(proteome$Gene.names, proteome$UniprotID, sep = "_")

# Create EList object
group <- as.factor(targets$Treatment)

proteome_Elist <- new("EList", list(
  E = as.matrix(proteome[sample_cols]), 
  genes = proteome[, 1:5],  # Gene annotation columns
  group = group[match(sample_cols, targets$Sample)]
))

cat("Initial EList dimensions:", dim(proteome_Elist$E), "\n")
```

```{r elist_observed}
# Convert zeros to NA
proteome_Elist$E[proteome_Elist$E == 0] <- NA

# Basic presence filter - keep proteins detected in at least 3 samples
is.exprs <- rowSums(!is.na(proteome_Elist$E)) >= 3
proteome_Elist$E <- proteome_Elist$E[is.exprs, ]
proteome_Elist$genes <- proteome_Elist$genes[is.exprs, ]

# Ensure rownames are preserved in genes
rownames(proteome_Elist$genes) <- rownames(proteome_Elist$E)

cat("After basic filtering:\n")
cat("  Proteins retained:", sum(is.exprs), "\n")
cat("  Proteins removed:", sum(!is.exprs), "\n")

# Missingness summary
cat("\nMissing values per sample:\n")
print(colSums(is.na(proteome_Elist$E)))
cat("\nProteins with >50% missing:", sum(rowMeans(is.na(proteome_Elist$E)) > 0.5), "\n")
```

### Filtering

```{r elist_filtering}
# NRG1 samples (columns 1-3: NRG124h_1_Disease, NRG124h_2_Disease, NRG124h_3_Disease)
is.exprs.nrg1 <- rowSums(!is.na(proteome_Elist$E[, 1:3]) & proteome_Elist$E[, 1:3] != 0, na.rm = TRUE) >= 2

# Scramble samples (columns 4-6: Scramble_1_Healthy, Scramble_2_Healthy, Scramble_3_Healthy)
is.exprs.scramble <- rowSums(!is.na(proteome_Elist$E[, 4:6]) & proteome_Elist$E[, 4:6] != 0, na.rm = TRUE) >= 2

# Keep proteins present in at least one condition
is.exprs <- is.exprs.nrg1 | is.exprs.scramble

cat("Filtering summary:\n")
cat("  Proteins in NRG1 (>=2/3 reps):", sum(is.exprs.nrg1), "\n")
cat("  Proteins in Scramble (>=2/3 reps):", sum(is.exprs.scramble), "\n")
cat("  Proteins retained (union):", sum(is.exprs), "\n")

proteome_Elist$E <- proteome_Elist$E[is.exprs, ]
proteome_Elist$genes <- proteome_Elist$genes[is.exprs, ]
proteome_Elist$group <- as.factor(targets$Treatment[match(colnames(proteome_Elist$E), targets$Sample)])

# Ensure rownames are preserved in genes
rownames(proteome_Elist$genes) <- rownames(proteome_Elist$E)
```

```{r filtering_plot, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
tmp <- data.frame(
  Sample = colnames(proteome_Elist$E), 
  Proportion = colMeans(!is.na(proteome_Elist$E)),
  Group = group[match(colnames(proteome_Elist$E), targets$Sample)]
)

tmp <- tmp[order(tmp$Group), ]
tmp$Sample <- factor(tmp$Sample, levels = tmp$Sample)

p <- ggplot(tmp, aes(x = Sample, y = Proportion, fill = Group)) + 
  geom_bar(stat = "identity") +
  scale_fill_manual(values = treatment_colors) +
  scale_y_continuous(expand = c(0, 0), limits = c(0, 1)) +
  labs(title = "Proportion of Features Present After Filtering",
       x = "Sample",
       y = "Proportion") +
  theme_classic() +
  theme(axis.text.x = element_text(hjust = 1, vjust = 0.5, angle = 90), 
        plot.title = element_text(hjust = 0.5))

print(p)
ggsave("results/qc/Proportion_Features_Filtered.svg", plot = p, width = 10, height = 6, device = "svg")
ggsave("results/qc/Proportion_Features_Filtered.jpeg", plot = p, width = 10, height = 6, device = "jpeg", dpi = 300)
```

```{r log2}
proteome_Elist_log2 <- proteome_Elist
proteome_Elist_log2$E <- log2(proteome_Elist$E)

cat("Raw intensities summary:\n")
print(summary(c(proteome_Elist$E)))
cat("\nLog2 intensities summary:\n")
print(summary(c(proteome_Elist_log2$E)))
cat("\nDimensions:", dim(proteome_Elist_log2), "\n")
```

```{r median_center_function}
median_center <- function(x) {
  centered <- apply(x, 2, function(y) {
    med <- median(y, na.rm = TRUE)
    y - med
  })
  return(centered)
}
```

```{r median_centering}
proteome_Elist_centered <- proteome_Elist_log2
proteome_Elist_centered$E <- median_center(proteome_Elist_log2$E)

# Report median centering values
centering_values <- apply(proteome_Elist_log2$E, 2, median, na.rm = TRUE)
cat("Median centering values applied:\n")
print(centering_values)
```

```{r log2_plot_post_processing, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
par(bty = "L")
plotDensities(proteome_Elist_centered$E, 
              group = droplevels(group[match(colnames(proteome_Elist_centered$E), targets$Sample)]),
              col = treatment_colors,
              legend = "topright",
              main = "Log2 Intensity Distribution (After Normalization)")
```

```{r RLE_plot_post_processing, fig.align="center", fig.width=10, fig.height=6, dpi=300, dev="png"}
sample_treatments <- targets$Treatment[match(colnames(proteome_Elist_centered$E), targets$Sample)]
plot_colors <- treatment_colors[sample_treatments]

boxplot(proteome_Elist_centered$E - rowMedians(as.matrix(proteome_Elist_centered$E), na.rm = TRUE), 
        col = plot_colors,
        las = 2,
        main = "RLE - Normalized Non-imputed Intensities")
abline(h = 0)
```

### MDS and PCA

```{r mds, fig.align="center", echo=FALSE, fig.width=10, fig.height=8, dpi=300, dev="png"}
treatments <- targets$Treatment[match(colnames(proteome_Elist_centered$E), targets$Sample)]

plotMDS(proteome_Elist_centered, 
       col = treatment_colors[treatments],
       pch = 16,
       cex = 2,
       main = "MDS After Normalization")

legend("topright", 
      legend = unique(treatments), 
      col = treatment_colors[unique(treatments)],
      pch = 16,
      bty = "n")
```

```{r PCA_plot_post_processing, fig.align="center", fig.width=10, fig.height=8, dpi=300, dev="png"}
# Perform PCA on complete cases
pca_data <- prcomp(t(na.omit(proteome_Elist_centered$E)), scale. = TRUE)

# Calculate variance explained
var_explained <- round(100 * pca_data$sdev^2 / sum(pca_data$sdev^2), 1)

# Prepare plot data
plotMe <- as.data.frame(pca_data$x[, 1:2])
plotMe <- rownames_to_column(plotMe, var = "Sample")
plotMe$Treatment <- targets$Treatment[match(plotMe$Sample, targets$Sample)]
plotMe$Condition <- targets$Condition[match(plotMe$Sample, targets$Sample)]

# Create PCA plot
p <- ggplot(plotMe, aes(x = PC1, y = PC2, color = Treatment, label = Sample)) + 
  geom_point(size = 4) +
  geom_text_repel(size = 3, max.overlaps = 10) +
  scale_color_manual(values = treatment_colors) +
  theme_classic() + 
  labs(title = "PCA: PC1 vs PC2",
       x = paste0("PC1 (", var_explained[1], "%)"),
       y = paste0("PC2 (", var_explained[2], "%)"))

print(p)
ggsave("results/qc/PCA_plot.svg", plot = p, width = 10, height = 8, device = "svg")
ggsave("results/qc/PCA_plot.jpeg", plot = p, width = 10, height = 8, device = "jpeg", dpi = 300)
```

```{r sample_correlation_heatmap, fig.align="center", fig.width=8, fig.height=8, dpi=300, dev="png"}
# Sample correlation heatmap to check replicate consistency
cor_matrix <- cor(proteome_Elist_centered$E, use = "pairwise.complete.obs")

# Annotation for samples
cor_anno <- data.frame(
  Treatment = targets$Treatment[match(colnames(cor_matrix), targets$Sample)]
)
rownames(cor_anno) <- colnames(cor_matrix)

cor_colors <- list(Treatment = treatment_colors)

# Correlation heatmap palette (blue to white to red)
cor_palette <- colorRampPalette(c("#0072B2", "white", "#D55E00"))(100)

pheatmap(cor_matrix,
         display_numbers = TRUE,
         number_format = "%.3f",
         fontsize_number = 9,
         annotation_col = cor_anno,
         annotation_row = cor_anno,
         annotation_colors = cor_colors,
         color = cor_palette,
         breaks = seq(min(cor_matrix, na.rm = TRUE), 1, length.out = 101),
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         fontsize = 10,
         border_color = NA,
         main = "Sample Correlation Heatmap",
         filename = "results/qc/Sample_Correlation_Heatmap.svg",
         width = 8, height = 8)

pheatmap(cor_matrix,
         display_numbers = TRUE,
         number_format = "%.3f",
         fontsize_number = 9,
         annotation_col = cor_anno,
         annotation_row = cor_anno,
         annotation_colors = cor_colors,
         color = cor_palette,
         breaks = seq(min(cor_matrix, na.rm = TRUE), 1, length.out = 101),
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         fontsize = 10,
         border_color = NA,
         main = "Sample Correlation Heatmap",
         filename = "results/qc/Sample_Correlation_Heatmap.jpeg",
         width = 8, height = 8)

# Display in notebook
pheatmap(cor_matrix,
         display_numbers = TRUE,
         number_format = "%.3f",
         fontsize_number = 9,
         annotation_col = cor_anno,
         annotation_row = cor_anno,
         annotation_colors = cor_colors,
         color = cor_palette,
         breaks = seq(min(cor_matrix, na.rm = TRUE), 1, length.out = 101),
         cluster_rows = TRUE,
         cluster_cols = TRUE,
         fontsize = 10,
         border_color = NA,
         main = "Sample Correlation Heatmap")
```

### Saving Processed Data

```{r saving_proteome_elist}
saveRDS(proteome_Elist_centered, file = "proteome_Elist_centered.rds")
cat("Processed data saved to: proteome_Elist_centered.rds\n")
```

### Differential Analysis

```{r limma_design}
# Simple design for NRG1 vs Scramble comparison
# Ensure Treatment is a factor
targets$Treatment <- factor(targets$Treatment, levels = c("Scramble", "NRG1"))

design <- model.matrix(~ 0 + Treatment, data = targets)
colnames(design) <- gsub("Treatment", "", colnames(design))

cat("Design matrix:\n")
print(design)
cat("\nSample distribution:\n")
print(table(targets$Treatment))
```

```{r limma_fit, fig.align="center", echo=FALSE, fig.width=10, fig.height=6, dpi=300, dev="png"}
fit_proteome <- lmFit(proteome_Elist_centered$E, design)
cat("Average expression summary:\n")
print(summary(fit_proteome$Amean))
```

```{r limma_contrasts}
# Single contrast: NRG1 vs Scramble
contrast_matrix <- makeContrasts(
  NRG1_vs_Scramble = NRG1 - Scramble,
  levels = design
)

cat("Contrast matrix:\n")
print(contrast_matrix)
```

```{r fitting_contrasts_ebayes}
fit_proteome <- contrasts.fit(fit_proteome, contrast_matrix)
fit_proteome <- eBayes(fit_proteome, trend = TRUE)
```

```{r limma_variance_plot, fig.align="center", echo=FALSE, fig.width=10, fig.height=6, dpi=300, dev="png"}
plotSA(fit_proteome, main = "Mean-Variance Trend - NRG1 vs Scramble Proteome")
```

```{r limma_results}
# Get differential expression results
DT_proteome <- decideTests(fit_proteome, p.value = 0.05, adjust.method = "BH")
cat("Summary of differential expression:\n")
print(summary(DT_proteome))

# Get top table results
tt_proteome <- topTable(fit_proteome, coef = "NRG1_vs_Scramble", n = Inf, sort.by = "P")

# Count significant proteins
sig_proteins <- sum(tt_proteome$adj.P.Val < 0.05, na.rm = TRUE)
cat("\nSignificant proteins (FDR < 0.05):", sig_proteins, "\n")
```

```{r de_proteins_function}
output_dir <- "results/differential"

extract_and_save_results <- function(fit, contrast_name, elist, output_dir) {
  # Extract statistical results
  results <- topTable(fit, coef = contrast_name, number = Inf, sort.by = "P")
  
  # Save original rownames as Gene.names column
  results$Gene.names <- rownames(results)
  
  # Add annotation columns
  if(!is.null(elist$genes) && nrow(elist$genes) > 0) {
    for(col in colnames(elist$genes)) {
      if(col %in% colnames(results)) next
      results[[col]] <- NA
    }
    
    common_ids <- intersect(rownames(results), rownames(elist$genes))
    
    if(length(common_ids) > 0) {
      for(col in colnames(elist$genes)) {
        if(!(col %in% colnames(results))) next
        results[common_ids, col] <- elist$genes[common_ids, col]
      }
    }
  }
  
  # Add classification columns
  results$Significant <- ifelse(results$adj.P.Val < 0.05, "FDR < 0.05", "Not Significant")
  results$Effect_Size <- ifelse(abs(results$logFC) > 1, "Large Effect (|logFC| > 1)", "Small Effect")
  results$Category <- paste(results$Significant, results$Effect_Size, sep = " & ")
  
  # Save to file
  filename <- file.path(output_dir, paste0(contrast_name, "_results.csv"))
  write.csv(results, filename, row.names = FALSE)
  
  cat(sprintf("Saved %s: %d features, %d significant (FDR < 0.05)\n", 
              contrast_name, nrow(results), sum(results$adj.P.Val < 0.05)))
  
  return(results)
}

# Extract results
all_results <- extract_and_save_results(
  fit = fit_proteome, 
  contrast_name = "NRG1_vs_Scramble",
  elist = proteome_Elist_centered,
  output_dir = output_dir
)
```

```{r volcano_plot_function}
regulation_colors <- c(
  "Upregulated" = "#14767F",      # RENEW Teal  
  "Downregulated" = "#D55E00",    # Oxygenated Red
  "Significant" = "#CC79A7",      # Cool Purple (significant but low FC)
  "Not Significant" = "#999999"   # Slate Grey
)

create_volcano_plot <- function(fit, contrast_name, elist, 
                                fdr_threshold = 0.05, 
                                fc_cutoff = 1,
                                n_label = 10) {
  
  # Extract results
  results <- topTable(fit, coef = contrast_name, number = Inf, sort.by = "none")
  
  # Get gene names from elist annotation using rownames for matching
  # Use drop=TRUE to ensure we get a vector, not a matrix/data.frame
  if(!is.null(rownames(elist$genes))) {
    results$Gene <- as.character(elist$genes[rownames(results), "Gene.names", drop = TRUE])
  } else {
    # Fallback: try to extract from rownames (format: GeneName_UniprotID)
    results$Gene <- sub("_[^_]+$", "", rownames(results))
  }
  
  # For proteins without gene names, use UniprotID as fallback
  missing_genes <- is.na(results$Gene) | results$Gene == "" | results$Gene == "NA"
  if(any(missing_genes)) {
    results$Gene[missing_genes] <- sub(".*_", "", rownames(results)[missing_genes])
  }
  
  # Classify proteins
  results$Regulation <- "Not Significant"
  results$Regulation[results$adj.P.Val < fdr_threshold] <- "Significant"
  results$Regulation[results$adj.P.Val < fdr_threshold & results$logFC > fc_cutoff] <- "Upregulated"
  results$Regulation[results$adj.P.Val < fdr_threshold & results$logFC < -fc_cutoff] <- "Downregulated"
  results$Regulation <- factor(results$Regulation, 
                               levels = c("Upregulated", "Downregulated", "Significant", "Not Significant"))
  
  # Count
  n_up <- sum(results$Regulation == "Upregulated")
  n_down <- sum(results$Regulation == "Downregulated")
  n_sig <- sum(results$Regulation == "Significant")
  
  # Select top genes to label
  results$Label <- ""
  
  # Top upregulated
  up_genes <- results[results$Regulation == "Upregulated", ]
  if(nrow(up_genes) > 0) {
    up_genes <- up_genes[order(-up_genes$logFC), ]
    n_up_label <- min(n_label, nrow(up_genes))
    results$Label[rownames(results) %in% rownames(up_genes)[1:n_up_label]] <- 
      results$Gene[rownames(results) %in% rownames(up_genes)[1:n_up_label]]
  }
  
  # Top downregulated
  down_genes <- results[results$Regulation == "Downregulated", ]
  if(nrow(down_genes) > 0) {
    down_genes <- down_genes[order(down_genes$logFC), ]
    n_down_label <- min(n_label, nrow(down_genes))
    results$Label[rownames(results) %in% rownames(down_genes)[1:n_down_label]] <- 
      results$Gene[rownames(results) %in% rownames(down_genes)[1:n_down_label]]
  }
  
  # Create volcano plot
  p <- ggplot(results, aes(x = logFC, y = -log10(adj.P.Val), color = Regulation)) +
    geom_point(alpha = 0.4, size = 2) +
    geom_text_repel(aes(label = Label), 
                   size = 4, 
                   max.overlaps = 20,
                   box.padding = 0.5,
                   point.padding = 0.5,
                   segment.color = "grey50",
                   segment.size = 0.5,
                   min.segment.length = 0,
                   show.legend = FALSE) +
    scale_color_manual(values = regulation_colors) +
    geom_hline(yintercept = -log10(fdr_threshold), linetype = "dashed", color = "black", linewidth = 0.8) +
    geom_vline(xintercept = c(-fc_cutoff, fc_cutoff), linetype = "dashed", color = "black", linewidth = 0.8) +
    labs(
      title = paste0("Volcano Plot: ", contrast_name),
      subtitle = paste0("Up: ", n_up, " | Down: ", n_down, " | Sig (low FC): ", n_sig),
      x = "log2 Fold Change",
      y = "-log10(FDR)",
      color = "Regulation"
    ) +
    theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
      plot.subtitle = element_text(hjust = 0.5, size = 14),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 12),
      legend.position = "right",
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    )
  
  return(p)
}
```

```{r proteome_volcano, warnings=FALSE, fig.align="center", echo=FALSE, fig.width=12, fig.height=8, dpi=300, dev="png"}
p <- create_volcano_plot(
  fit = fit_proteome,
  contrast_name = "NRG1_vs_Scramble",
  elist = proteome_Elist_centered,
  fdr_threshold = 0.05,
  fc_cutoff = 1,
  n_label = 15
)

print(p)
ggsave("results/volcano/NRG1_vs_Scramble_volcano_FDR.svg", plot = p, width = 12, height = 8, device = "svg")
ggsave("results/volcano/NRG1_vs_Scramble_volcano_FDR.jpeg", plot = p, width = 12, height = 8, device = "jpeg", dpi = 300)
```

```{r volcano_plot_pvalue_function}
# Volcano plot function using p-value instead of FDR
create_volcano_plot_pvalue <- function(fit, contrast_name, elist, 
                                       pval_threshold = 0.05, 
                                       fc_cutoff = 1,
                                       n_label = 10) {
  
  # Extract results
  results <- topTable(fit, coef = contrast_name, number = Inf, sort.by = "none")
  
  # Get gene names from elist annotation using rownames for matching
  if(!is.null(rownames(elist$genes))) {
    results$Gene <- as.character(elist$genes[rownames(results), "Gene.names", drop = TRUE])
  } else {
    results$Gene <- sub("_[^_]+$", "", rownames(results))
  }
  
  # For proteins without gene names, use UniprotID as fallback
  missing_genes <- is.na(results$Gene) | results$Gene == "" | results$Gene == "NA"
  if(any(missing_genes)) {
    results$Gene[missing_genes] <- sub(".*_", "", rownames(results)[missing_genes])
  }
  
  # Classify proteins using P.Value instead of adj.P.Val
  results$Regulation <- "Not Significant"
  results$Regulation[results$P.Value < pval_threshold] <- "Significant"
  results$Regulation[results$P.Value < pval_threshold & results$logFC > fc_cutoff] <- "Upregulated"
  results$Regulation[results$P.Value < pval_threshold & results$logFC < -fc_cutoff] <- "Downregulated"
  results$Regulation <- factor(results$Regulation, 
                               levels = c("Upregulated", "Downregulated", "Significant", "Not Significant"))
  
  # Count
  n_up <- sum(results$Regulation == "Upregulated")
  n_down <- sum(results$Regulation == "Downregulated")
  n_sig <- sum(results$Regulation == "Significant")
  
  # Select top genes to label
  results$Label <- ""
  
  # Top upregulated
  up_genes <- results[results$Regulation == "Upregulated", ]
  if(nrow(up_genes) > 0) {
    up_genes <- up_genes[order(-up_genes$logFC), ]
    n_up_label <- min(n_label, nrow(up_genes))
    results$Label[rownames(results) %in% rownames(up_genes)[1:n_up_label]] <- 
      results$Gene[rownames(results) %in% rownames(up_genes)[1:n_up_label]]
  }
  
  # Top downregulated
  down_genes <- results[results$Regulation == "Downregulated", ]
  if(nrow(down_genes) > 0) {
    down_genes <- down_genes[order(down_genes$logFC), ]
    n_down_label <- min(n_label, nrow(down_genes))
    results$Label[rownames(results) %in% rownames(down_genes)[1:n_down_label]] <- 
      results$Gene[rownames(results) %in% rownames(down_genes)[1:n_down_label]]
  }
  
  # Create volcano plot
  p <- ggplot(results, aes(x = logFC, y = -log10(P.Value), color = Regulation)) +
    geom_point(alpha = 0.4, size = 2) +
    geom_text_repel(aes(label = Label), 
                   size = 4, 
                   max.overlaps = 20,
                   box.padding = 0.5,
                   point.padding = 0.5,
                   segment.color = "grey50",
                   segment.size = 0.5,
                   min.segment.length = 0,
                   show.legend = FALSE) +
    scale_color_manual(values = regulation_colors) +
    geom_hline(yintercept = -log10(pval_threshold), linetype = "dashed", color = "black", linewidth = 0.8) +
    geom_vline(xintercept = c(-fc_cutoff, fc_cutoff), linetype = "dashed", color = "black", linewidth = 0.8) +
    labs(
      title = paste0("Volcano Plot (p-value): ", contrast_name),
      subtitle = paste0("Up: ", n_up, " | Down: ", n_down, " | Sig (low FC): ", n_sig),
      x = "log2 Fold Change",
      y = "-log10(p-value)",
      color = "Regulation"
    ) +
    theme_classic(base_size = 14) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", size = 18),
      plot.subtitle = element_text(hjust = 0.5, size = 14),
      axis.title = element_text(size = 16),
      axis.text = element_text(size = 12),
      legend.position = "right",
      legend.title = element_text(size = 14),
      legend.text = element_text(size = 12)
    )
  
  return(p)
}
```

```{r proteome_volcano_pvalue, warnings=FALSE, fig.align="center", echo=FALSE, fig.width=12, fig.height=8, dpi=300, dev="png"}
p_pval <- create_volcano_plot_pvalue(
  fit = fit_proteome,
  contrast_name = "NRG1_vs_Scramble",
  elist = proteome_Elist_centered,
  pval_threshold = 0.05,
  fc_cutoff = 1,
  n_label = 15
)

print(p_pval)
ggsave("results/volcano/NRG1_vs_Scramble_volcano_pvalue.svg", plot = p_pval, width = 12, height = 8, device = "svg")
ggsave("results/volcano/NRG1_vs_Scramble_volcano_pvalue.jpeg", plot = p_pval, width = 12, height = 8, device = "jpeg", dpi = 300)
```

### Heatmap of Significant Proteins

```{r proteome_heatmap, fig.align="center", echo=FALSE, fig.width=10, fig.height=10, dpi=300, dev="png"}
# Extract significant proteins
sig_proteins_df <- all_results[all_results$adj.P.Val < 0.05, ]

if(nrow(sig_proteins_df) > 0) {
  # Get expression data for significant proteins
  sig_features <- rownames(sig_proteins_df)
  sig_rows <- rownames(proteome_Elist_centered$E) %in% sig_features
  sig_data <- proteome_Elist_centered$E[sig_rows, ]
  
  # Convert to matrix
  if(is.data.frame(sig_data)) {
    sig_data <- as.matrix(sig_data)
  }
  
  # Clean data
  sig_data[!is.finite(sig_data)] <- NA
  sig_data <- na.omit(sig_data)
  
  if(nrow(sig_data) > 1) {
    # Scale data row-wise
    sig_data_scaled <- t(scale(t(sig_data)))
    sig_data_scaled[sig_data_scaled > 3] <- 3
    sig_data_scaled[sig_data_scaled < -3] <- -3
    
    # Annotation
    HManno_col <- data.frame(
      Treatment = targets$Treatment[match(colnames(sig_data_scaled), targets$Sample)]
    )
    rownames(HManno_col) <- colnames(sig_data_scaled)
    
    HManno_colors <- list(
      Treatment = treatment_colors
    )
    
    # Palette
    heatmap_palette <- colorRampPalette(c(
      "#301934",    # Very dark purple
      "#882255",    # Deep burgundy
      "#8B4F9F",    # Medium purple
      "#F1C232",    # Light gold
      "#F0E68C"     # Khaki
    ))(100)
    
    # Generate heatmap and save
    pheatmap(sig_data_scaled, 
             scale = "none",  
             show_rownames = nrow(sig_data_scaled) <= 50,
             annotation_col = HManno_col,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             clustering_distance_rows = "correlation",
             clustering_method = "complete",
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Significant Proteins (FDR < 0.05): n = ", nrow(sig_data_scaled)),
             filename = "results/differential/Significant_Proteins_Heatmap.svg",
             width = 10, height = 10)
    
    pheatmap(sig_data_scaled, 
             scale = "none",  
             show_rownames = nrow(sig_data_scaled) <= 50,
             annotation_col = HManno_col,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             clustering_distance_rows = "correlation",
             clustering_method = "complete",
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Significant Proteins (FDR < 0.05): n = ", nrow(sig_data_scaled)),
             filename = "results/differential/Significant_Proteins_Heatmap.jpeg",
             width = 10, height = 10)
    
    # Display in notebook
    pheatmap(sig_data_scaled, 
             scale = "none",  
             show_rownames = nrow(sig_data_scaled) <= 50,
             annotation_col = HManno_col,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             clustering_distance_rows = "correlation",
             clustering_method = "complete",
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Significant Proteins (FDR < 0.05): n = ", nrow(sig_data_scaled)))
  } else {
    cat("Not enough significant proteins for heatmap\n")
  }
} else {
  cat("No significant proteins found at FDR < 0.05\n")
}
```

### Heatmap of Top Differentially Expressed Proteins

```{r top_de_heatmap, fig.align="center", echo=FALSE, fig.width=12, fig.height=10, dpi=300, dev="png"}
# Select top N upregulated and downregulated proteins
n_top <- 25  # Number of top proteins per direction

# Get top upregulated (highest logFC with significant p-value)
top_up <- all_results %>%
  filter(P.Value < 0.05, logFC > 0) %>%
  arrange(desc(logFC)) %>%
  head(n_top)

# Get top downregulated (lowest logFC with significant p-value)
top_down <- all_results %>%
  filter(P.Value < 0.05, logFC < 0) %>%
  arrange(logFC) %>%
  head(n_top)

# Combine
top_de <- rbind(top_up, top_down)

if(nrow(top_de) > 1) {
  # Get expression data
  top_features <- rownames(top_de)
  top_rows <- rownames(proteome_Elist_centered$E) %in% top_features
  top_data <- proteome_Elist_centered$E[top_rows, ]
  
  # Convert to matrix
  if(is.data.frame(top_data)) {
    top_data <- as.matrix(top_data)
  }
  
  # Clean data
  top_data[!is.finite(top_data)] <- NA
  top_data <- na.omit(top_data)
  
  if(nrow(top_data) > 1) {
    # Scale data row-wise
    top_data_scaled <- t(scale(t(top_data)))
    top_data_scaled[top_data_scaled > 3] <- 3
    top_data_scaled[top_data_scaled < -3] <- -3
    
    # Get gene names for row labels
    row_genes <- proteome_Elist_centered$genes[rownames(top_data_scaled), "Gene.names", drop = TRUE]
    row_genes[is.na(row_genes) | row_genes == ""] <- sub(".*_", "", rownames(top_data_scaled)[is.na(row_genes) | row_genes == ""])
    rownames(top_data_scaled) <- row_genes
    
    # Create row annotation for regulation direction
    row_anno <- data.frame(
      Regulation = ifelse(rownames(top_data) %in% rownames(top_up), "Upregulated", "Downregulated")
    )
    rownames(row_anno) <- row_genes
    
    # Annotation
    HManno_col <- data.frame(
      Treatment = targets$Treatment[match(colnames(top_data_scaled), targets$Sample)]
    )
    rownames(HManno_col) <- colnames(top_data_scaled)
    
    HManno_colors <- list(
      Treatment = treatment_colors,
      Regulation = c("Upregulated" = "#14767F", "Downregulated" = "#D55E00")
    )
    
    # Palette
    heatmap_palette <- colorRampPalette(c(
      "#301934",    # Very dark purple
      "#882255",    # Deep burgundy
      "#8B4F9F",    # Medium purple
      "#F1C232",    # Light gold
      "#F0E68C"     # Khaki
    ))(100)
    
    # Save heatmaps
    pheatmap(top_data_scaled, 
             scale = "none",  
             show_rownames = TRUE,
             annotation_col = HManno_col,
             annotation_row = row_anno,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             cluster_rows = FALSE,  # Keep up/down grouped
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Top ", n_top, " Up/Downregulated Proteins (p < 0.05)"),
             filename = "results/differential/Top_DE_Proteins_Heatmap.svg",
             width = 12, height = 10)
    
    pheatmap(top_data_scaled, 
             scale = "none",  
             show_rownames = TRUE,
             annotation_col = HManno_col,
             annotation_row = row_anno,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             cluster_rows = FALSE,
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Top ", n_top, " Up/Downregulated Proteins (p < 0.05)"),
             filename = "results/differential/Top_DE_Proteins_Heatmap.jpeg",
             width = 12, height = 10)
    
    # Display in notebook
    pheatmap(top_data_scaled, 
             scale = "none",  
             show_rownames = TRUE,
             annotation_col = HManno_col,
             annotation_row = row_anno,
             annotation_colors = HManno_colors,
             color = heatmap_palette,
             cluster_cols = TRUE,
             cluster_rows = FALSE,
             fontsize = 10,
             fontsize_row = 8,
             border_color = NA,
             main = paste0("Top ", n_top, " Up/Downregulated Proteins (p < 0.05)"))
  } else {
    cat("Not enough proteins for top DE heatmap\n")
  }
} else {
  cat("Not enough differentially expressed proteins for heatmap\n")
}
```

### Gene Ontology Enrichment Analysis

```{r go_enrichment, fig.align="center", fig.width=12, fig.height=10, dpi=300, dev="png"}

# Create GO folder
dir.create("results/GO", recursive = TRUE, showWarnings = FALSE)

# Get significant genes
sig_up <- all_results$Gene.names[all_results$P.Value < 0.05 & all_results$logFC > 1]
sig_down <- all_results$Gene.names[all_results$P.Value < 0.05 & all_results$logFC < -1]

# Remove NA
sig_up <- sig_up[!is.na(sig_up)]
sig_down <- sig_down[!is.na(sig_down)]

cat("Upregulated genes:", length(sig_up), "\n")
cat("Downregulated genes:", length(sig_down), "\n")

# Run enrichment if enough genes
if(length(sig_up) >= 5) {
  cat("\n--- Upregulated GO Enrichment ---\n")
  ego_up <- enrichGO(gene = sig_up,
                     OrgDb = org.Hs.eg.db,
                     keyType = "SYMBOL",
                     ont = "BP",
                     pAdjustMethod = "BH",
                     pvalueCutoff = 0.05)
  
  if(!is.null(ego_up) && nrow(ego_up@result) > 0) {
    p_go_up <- dotplot(ego_up, showCategory = 15, title = "GO BP - Upregulated in NRG1")
    print(p_go_up)
    
    # Save
    ggsave("results/GO/GO_BP_Upregulated.svg", plot = p_go_up, width = 12, height = 10, device = "svg")
    ggsave("results/GO/GO_BP_Upregulated.pdf", plot = p_go_up, width = 12, height = 10, device = "pdf")
  }
}

if(length(sig_down) >= 5) {
  cat("\n--- Downregulated GO Enrichment ---\n")
  ego_down <- enrichGO(gene = sig_down,
                       OrgDb = org.Hs.eg.db,
                       keyType = "SYMBOL",
                       ont = "BP",
                       pAdjustMethod = "BH",
                       pvalueCutoff = 0.05)
  
  if(!is.null(ego_down) && nrow(ego_down@result) > 0) {
    p_go_down <- dotplot(ego_down, showCategory = 15, title = "GO BP - Upregulated in Healthy")
    print(p_go_down)
    
    # Save
    ggsave("results/GO/GO_BP_Downregulated.svg", plot = p_go_down, width = 12, height = 10, device = "svg")
    ggsave("results/GO/GO_BP_Downregulated.pdf", plot = p_go_down, width = 12, height = 10, device = "pdf")
  }
}

# Check saved files
cat("\nSaved GO plots:\n")
list.files("results/GO/")
```

Key changes:
- Created `results_24h/GO/` folder
- Stored plots in variables (`p_go_up`, `p_go_down`)
- Added `ggsave()` for both SVG and PDF formats
- Listed saved files at the end
```

### Summary Statistics

```{r summary_stats}
cat("=== Analysis Summary ===\n\n")
cat("Input data:\n")
cat("  - Total proteins in dataset:", nrow(proteomics_data), "\n")
cat("  - Proteins after filtering:", nrow(proteome_Elist_centered$E), "\n")

cat("\nDifferential expression (NRG1 vs Scramble):\n")
cat("  - Upregulated (FDR < 0.05, logFC > 1):", sum(all_results$adj.P.Val < 0.05 & all_results$logFC > 1), "\n")
cat("  - Downregulated (FDR < 0.05, logFC < -1):", sum(all_results$adj.P.Val < 0.05 & all_results$logFC < -1), "\n")
cat("  - Significant (FDR < 0.05, any FC):", sum(all_results$adj.P.Val < 0.05), "\n")

cat("\nTop 10 upregulated proteins:\n")
top_up <- head(all_results[order(-all_results$logFC), c("Gene.names", "logFC", "adj.P.Val")], 10)
print(top_up)

cat("\nTop 10 downregulated proteins:\n")
top_down <- head(all_results[order(all_results$logFC), c("Gene.names", "logFC", "adj.P.Val")], 10)
print(top_down)
```

### ERBB Analysis

```{r ERBB}
dir.create("results/differential", recursive = TRUE, showWarnings = FALSE)


library(dplyr)
library(ggplot2)

erbb_genes <- c(
# Receptors
  "EGFR", "ERBB2", "ERBB3", "ERBB4",
  # Ligands
  "EGF", "TGFA", "AREG", "BTC", "EREG", "HBEGF", "NRG1", "NRG2", "NRG3", "NRG4",
  # Adaptors & signaling
  "SHC1", "SHC2", "SHC3", "SHC4", "GRB2", "SOS1", "SOS2", "GAB1",
  "CRK", "CRKL", "NCK1", "NCK2", "CBL", "CBLB", "CBLC",
  # RAS-MAPK pathway
  "HRAS", "KRAS", "NRAS", "RAF1", "ARAF", "BRAF",
  "MAP2K1", "MAP2K2", "MAPK1", "MAPK3", "MAPK8", "MAPK9", "MAPK10",
  # PI3K-AKT pathway
  "PIK3CA", "PIK3CB", "PIK3CD", "PIK3CG", "PIK3R1", "PIK3R2", "PIK3R3",
  "AKT1", "AKT2", "AKT3", "MTOR", "GSK3B", "BAD",
  # Transcription factors & cell cycle
  "MYC", "ELK1", "JUN", "CDKN1A", "CDKN1B",
  # Other kinases
  "SRC", "ABL1", "ABL2", "PTK2", "PAK1", "PAK2", "PAK3", "PAK4",
  # STAT pathway
  "STAT5A", "STAT5B",
  # PLC pathway
  "PLCG1", "PLCG2", "PRKCA", "PRKCB", "PRKCG",
  # Calcium signaling
  "CAMK2A", "CAMK2B", "CAMK2D", "CAMK2G"
)

# Filter all_results for ERBB pathway genes
erbb_results <- all_results %>%
  filter(Gene.names %in% erbb_genes) %>%
  mutate(
    Significance = case_when(
      P.Value < 0.001 ~ "p < 0.001",
      P.Value < 0.01 ~ "p < 0.01",
      P.Value < 0.05 ~ "p < 0.05",
      TRUE ~ "NS"
    ),
    Regulation = case_when(
      logFC > 0 & P.Value < 0.05 ~ "Up",
      logFC < 0 & P.Value < 0.05 ~ "Down",
      TRUE ~ "NS"
    )
  ) %>%
  arrange(desc(logFC))

cat("ERBB pathway genes found in dataset:", nrow(erbb_results), "\n")
cat("Significant (p < 0.05):", sum(erbb_results$P.Value < 0.05), "\n")

# Create dot plot - only significant genes (p <= 0.05)
if(nrow(erbb_results) > 0) {
  
  # If there are duplicates, keep the one with lowest p-value
  # Then filter for significant only
  erbb_results_unique <- erbb_results %>%
    group_by(Gene.names) %>%
    slice_min(P.Value, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    filter(P.Value <= 0.05) %>%
    arrange(logFC)
  
  cat("Significant ERBB pathway genes:", nrow(erbb_results_unique), "\n")
  
  if(nrow(erbb_results_unique) > 0) {
    
    # Order by logFC
    erbb_results_unique$Gene.names <- factor(erbb_results_unique$Gene.names,    levels = erbb_results_unique$Gene.names)
    
    p_erbb <- ggplot(erbb_results_unique, aes(x = logFC, y = Gene.names)) +
      geom_point(aes(size = -log10(P.Value), color = logFC)) +
      geom_vline(xintercept = 0, linetype = "dashed", color = "grey50") +
      scale_color_gradient2(low = "#D55E00", mid = "grey80", high = "#14767F", midpoint = 0,
                            name = "log2FC") +
      scale_size_continuous(name = "-log10(p-value)", range = c(3, 8)) +
      labs(
        title = "ERBB Signaling Pathway - NRG1 vs Scramble",
        subtitle = paste0("Significant genes (p ≤ 0.05): ", nrow(erbb_results_unique)),
        x = "log2 Fold Change",
        y = ""
      ) +
      theme_classic(base_size = 12) +
      theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.text.y = element_text(size = 10),
        legend.position = "right"
      )
    
    print(p_erbb)
    
    # Save
    ggsave("results/differential/ERBB_pathway_dotplot.svg", plot = p_erbb, width = 10, height = max(6, nrow(erbb_results_unique) * 0.3), device = "svg")
    ggsave("results/differential/ERBB_pathway_dotplot.jpeg", plot = p_erbb, width = 10, height = max(6, nrow(erbb_results_unique) * 0.3), device = "jpeg", dpi = 300)
    
  } else {
    cat("No significant ERBB pathway genes found (p <= 0.05)\n")
  }
  
} else {
  cat("No ERBB pathway genes found in results\n")
}









# Create dot plot - single column, all genes aligned vertically
if(nrow(erbb_results) > 0) {
  
  # If there are duplicates, keep the one with lowest p-value
  # Then filter for significant only
  erbb_results_unique <- erbb_results %>%
    group_by(Gene.names) %>%
    slice_min(P.Value, n = 1, with_ties = FALSE) %>%
    ungroup() %>%
    filter(P.Value <= 0.05) %>%
    arrange(logFC)
  
  cat("Significant ERBB pathway genes:", nrow(erbb_results_unique), "\n")
  
  if(nrow(erbb_results_unique) > 0) {
    
    # Order by logFC
    erbb_results_unique$Gene.names <- factor(erbb_results_unique$Gene.names, 
                                             levels = erbb_results_unique$Gene.names)
    
    # Add a single x-axis category
    erbb_results_unique$Timepoint <- "24h"
    
    p_erbb <- ggplot(erbb_results_unique, aes(x = Timepoint, y = Gene.names)) +
      geom_point(aes(size = -log10(P.Value), color = logFC)) +
      scale_color_gradient2(low = "#D55E00", mid = "grey80", high = "#14767F", midpoint = 0,
                            name = "log2FC") +
      scale_size_continuous(name = "-log10(p-value)", range = c(3, 8)) +
      labs(
        title = "ERBB Signaling Pathway - NRG1 vs Scramble",
        subtitle = paste0("Significant genes (p ≤ 0.05): ", nrow(erbb_results_unique)),
        x = "",
        y = ""
      ) +
      theme_classic(base_size = 12) +
      theme(
        plot.title = element_text(hjust = 0.5, face = "bold", size = 14),
        plot.subtitle = element_text(hjust = 0.5, size = 11),
        axis.text.y = element_text(size = 10),
        axis.text.x = element_text(size = 12, face = "bold"),
        legend.position = "right"
      )
    
    print(p_erbb)
    
    # Save
    ggsave("results/differential/ERBB_pathway_dotplot.svg", plot = p_erbb, width = 6, height = max(6, nrow(erbb_results_unique) * 0.3), device = "svg")
    ggsave("results/differential/ERBB_pathway_dotplot.jpeg", plot = p_erbb, width = 6, height = max(6, nrow(erbb_results_unique) * 0.3), device = "jpeg", dpi = 300)
    
  } else {
    cat("No significant ERBB pathway genes found (p <= 0.05)\n")
  }
  
} else {
  cat("No ERBB pathway genes found in results\n")
}
```

### Session Info

```{r session_info}
sessionInfo()
```

